---
title: "transmission_experiment2"
author: "Tatum Katz"
date: "5/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = '~/Google Drive/fruit fly vector project')
```

# libraries

```{r}

library(ggplot2)
library(car)
library(tidyr)
library(dplyr)
library(kableExtra)
library(ggmosaic)
library(ggpubr)
library(stringr)
library(latex2exp)

```


# datasets

```{r}

one <- read.csv("clean data/transmission2_plate1.csv")
two <- read.csv("clean data/transmission2_plate2.csv")
three <- read.csv("clean data/transmission2_plate3.csv")

# remove water samples from plate three
three <- subset(three, str_length(Sample.Name)>=7)

# stack data

one <- rbind(one,two,three)

```

# more cleaning

```{r}

# remove qiagen negative and blanks
one <- subset(one, Sample.Name!="quiagen_neg" & Sample.Name!="blank")

# remove inhibited samples
one <- subset(one, inhibited!="1")

# remove "original" samples that were inhibited but aren't but keep the "rerun" versions
one <- subset(one, !grepl("*_o", one$Sample.Name))

# make new names for treatments
temp <- strsplit(one$Sample.Name, "_")
n.flies <- c()
prop.inf <- c()
fly.id <- c()
for(i in 1:length(temp)){
  if(length(temp[[i]])==3){
    n.flies[i] <- temp[[i]][1]
    prop.inf[i] <- temp[[i]][2]
    fly.id[i] <- temp[[i]][3]
  }else if(length(temp[[i]]==4)){
    n.flies[i] <- temp[[i]][1]
    prop.inf[i] <- temp[[i]][2]
    fly.id[i] <- paste0(temp[[i]][3], temp[[i]][4])
  }
}
one$fly.id <- fly.id
one$n.flies <- as.numeric(n.flies)
one$prop.inf <- as.numeric(prop.inf)

inf <- c()
for(i in 1:nrow(one)){
  if(one$Qty[i]>0){inf[i] <- 1}
  else if(one$Qty[i]==0){inf[i] <- 0}
}
one$inf <- factor(inf)

# note if you want to analyze loads, ones with r need to be calculated differently
# took 140/200 ul dna, respun and eluted into only 50ul

one <- na.omit(one) #removes negative treatments

```

# EDA - real data

```{r}

ggplot(one, aes(y=log(Qty*40+1), fill=factor(as.character(n.flies), ordered=T, levels=c("2","4","6","8","10","12","14","16")), x = as.character(prop.inf)))+
  geom_boxplot()+
  labs(x="Proportion Infected Flies", y="Total Zoospore Equivalents per Fly (log plus 1)",
       fill="Number of Flies")

```


```{r}

counts <- one %>%
  filter(prop.inf>10) %>%
  group_by(prop.inf, n.flies) %>%
  summarize(n=n(),
            n.inf=sum(as.numeric(as.character(inf)))) %>%
  mutate(assayed.inf = n.inf/n, # this will tell us how many got infected after removing those who should be infected
         sus.inf = n.inf-prop.inf*.01*n.flies) 


# make dict for facet labels
labs <- c("12.5"="12.5% Infecteds", "25"="25% Infecteds", 
          "50"="50% Infecteds", "24"="24% Infecteds")

# remove negatives and add label column from counts
one <- one %>%
  filter(prop.inf>10)

#  add label column from counts
barlabs <- c()
for(i in 1:nrow(one)){
  if(one$prop.inf[i]==12.5 & one$n.flies[i]==8){
    barlabs[i]<-counts$assayed.inf[1]
  }else if(one$prop.inf[i]==12.5 & one$n.flies[i]==16){
      barlabs[i]<-counts$assayed.inf[2]
  }else if(one$prop.inf[i]==24 & one$n.flies[i]==34){
    barlabs[i]<-counts$assayed.inf[3]
  }else if(one$prop.inf[i]==25 & one$n.flies[i]==4){
      barlabs[i]<-counts$assayed.inf[4]
  }else if(one$prop.inf[i]==25 & one$n.flies[i]==8){
      barlabs[i]<-counts$assayed.inf[5]
  }else if(one$prop.inf[i]==25 & one$n.flies[i]==12){
      barlabs[i]<-counts$assayed.inf[6]
  }else if(one$prop.inf[i]==25 & one$n.flies[i]==16){
      barlabs[i]<-counts$assayed.inf[7]
  }else if(one$prop.inf[i]==25 & one$n.flies[i]==34){
      barlabs[i]<-counts$assayed.inf[8]
  }else if(one$prop.inf[i]==50 & one$n.flies[i]==2){
      barlabs[i]<-counts$assayed.inf[9]
  }else if(one$prop.inf[i]==50 & one$n.flies[i]==4){
      barlabs[i]<-counts$assayed.inf[10]
  }else if(one$prop.inf[i]==50 & one$n.flies[i]==6){
      barlabs[i]<-counts$assayed.inf[11]
  }else if(one$prop.inf[i]==50 & one$n.flies[i]==8){
      barlabs[i]<-counts$assayed.inf[12]
  }else if(one$prop.inf[i]==50 & one$n.flies[i]==10){
      barlabs[i]<-counts$assayed.inf[13]
  }else if(one$prop.inf[i]==50 & one$n.flies[i]==12){
      barlabs[i]<-counts$assayed.inf[14]
  }else if(one$prop.inf[i]==50 & one$n.flies[i]==14){
      barlabs[i]<-counts$assayed.inf[15]
  }else if(one$prop.inf[i]==50 & one$n.flies[i]==16){
      barlabs[i]<-counts$assayed.inf[16]
  }else if(one$prop.inf[i]==50 & one$n.flies[i]==18){
      barlabs[i]<-counts$assayed.inf[17]
  }else if(one$prop.inf[i]==50 & one$n.flies[i]==32){
      barlabs[i]<-counts$assayed.inf[18]
  }
}


ggplot(one, aes(x=n.flies, fill=inf))+
  geom_bar(position="stack", stat="count")+
  facet_wrap(.~as.character(prop.inf),
             labeller = as_labeller(labs))+
  scale_x_continuous(breaks=c(2,4,6,8,10,12,14,16,18,32,34))+
  guides(color="none")+
  labs(x="Density of Flies", y="Number of Flies Infected",
       fill="Infected?")+
  geom_text(aes(label = round(barlabs,2), y=10), angle=45,
            position = position_dodge(width = 0.9),
            check_overlap = T, size=2)+
  scale_fill_discrete(limits=rev, labels=c("Yes","No"))
# 
# 
# ggplot(counts, aes(x=n.flies, y=assayed.inf, fill=factor(prop.inf)))+
#   geom_bar(stat="identity")+
#   facet_wrap(.~as.character(prop.inf),
#              labeller = as_labeller(labs))+
#   scale_x_continuous(breaks=seq(2,32,2))+
#   guides(color="none", fill="none")+
#   labs(x="Density of Flies", y="Final Proportion Infected",
#        fill="")
# 
#   geom_text(aes(label = round(barlabs,2), y=10), angle=45,
#             position = position_dodge(width = 0.9),
#             check_overlap = T)

```

# analysis: following rachowitz and briggs & tompros 2021

## first: write functions for each transmission function

thank you Amy Greer! (see matlab code)

## next: for each experiment, identify starting s, starting i, and ending i

```{r}

params <- counts

colnames(params) <- c("prop.inf", "start.n", "end.n", "end.inf", "assayed.inf", "sus.inf")
params$end.sus <- params$end.n-params$end.inf
params$start.inf <- params$prop.inf/100*params$start.n 
params$start.sus <- params$start.n - params$start.inf

dat <- data.frame(start.n=params$start.n,
                  start.sus=params$start.sus,
                  start.inf=params$start.inf,
                  end.n=params$end.n,
                  end.sus=params$end.sus,
                  end.inf=params$end.inf)

write.csv(dat, "matlab transmission fn fitting/trans_dat.csv", row.names=F)

```

## next:use a simplex search method  to find the values of the parameters that minimize the negative log likelihood of the data from each experiment, given the model.

thank you Amy Greer! see matlab code

# matlab results

```{r}

aicCR <- 75.75
aicFD <- 97.11
aicDD <- 63.51 # Greer calls this mass action (MA)
aicPwr <- 61.94
alpha_Pwr <- 2.16
beta_CR <- 0.097
beta_FD <- 1.00
beta_DD <- 0.015
beta_Pwr <- 0.00
caicCR <- 75.93
caic_FD <- 97.29
caic_DD <- 63.70
caic_Pwr <- 62.51

```













